GBM recurrence - Shot gun analysis of proteolytic products

### 2023-10-24 reports/html_gbm_recurrence_general_proteomics_manuscript_report.qmd
revision 2 commit

---
title: "GBM recurrence - Shotgun proteomics analysis"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---

@@

```{r fig.width=5.5, fig.height=4.5, warning=FALSE}
ggplot(data = plsda_variates,
       aes(x = comp1, 
           y = comp2)) +
  geom_point(size = 4, aes(shape = mixture, 
                           color = recurrence)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 2)) +
  labs(title = "sPLS-DA plot of samples - pat6 excluded \nBased on the abundance values of proteins") +
  xlab(paste("sPLS-DA Comp-1", 
             round(plsda_restwo6$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("sPLS-DA Comp-2", 
             round(plsda_restwo6$prop_expl_var$X[2]*100), 
             "% var explained")) + 
  ggforce::geom_mark_ellipse(mapping = aes(color = recurrence)) +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

@@

```{r fig.width=8, fig.height=6, echo=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05,
             Gene %in% c("ASAH1", "GPNMB", "SYNM", "CD14"))

@@

slim_data3 <- slim_data2 %>% 
  mutate(Abs_diff = rec-prim) %>%
  mutate(rank = row_number(Abs_diff)) %>% 
  dplyr::rename(Initial = prim, 
                Recurrent = rec) %>%
  ungroup() %>%
  filter(Protein %in% c("Q13510", 
                        "Q14956", 
                        "O15061",
                        "P05164")) %>%
  mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1",
                          Protein == "Q14956" ~ "GPNMB",
                          Protein == "O15061" ~ "SYNM",
                          Protein == "P05164" ~ "MPO"))
```

>>>>

---
title: "GBM recurrence - Shotgun proteomics analysis"
author: "Miguel Cosenza-Contreras"
format:
  html: 
    toc-depth: 4 # default is 3
    toc-title: Contents
    toc-location: left # default is right, option is 'body'
    number-sections: true
    number-depth: 3
    embed-resources: true
always_allow_html: true
editor: source
---

@@

```{r fig.width=5.5, fig.height=4.5, warning=FALSE}
ggplot(data = plsda_variates,
       aes(x = comp1, 
           y = comp2)) +
  geom_point(size = 4, aes(shape = mixture, 
                           color = recurrence)) +  
  geom_text(aes(label = patient), 
            position = position_nudge(x = 2)) +
  #labs(title = "sPLS-DA plot of samples - pat6 excluded \nBased on the abundance values of proteins") +
  xlab(paste("sPLS-DA Comp-1", 
             round(plsda_restwo6$prop_expl_var$X[1]*100), 
             "% var explained")) + 
  ylab(paste("sPLS-DA Comp-2", 
             round(plsda_restwo6$prop_expl_var$X[2]*100), 
             "% var explained")) + 
  ggforce::geom_mark_ellipse(mapping = aes(color = recurrence)) +
  theme(axis.text.x = element_text(hjust = 0.5, 
                                   vjust = 0, 
                                   size = 10, 
                                   angle = 360),
        axis.text.y = element_text(hjust = 0.95, 
                                   vjust = 0.2, 
                                   size = 10),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.key.height= unit(3, 'mm'),
        legend.key.width= unit(3, 'mm'),
        legend.position="bottom")
```

@@

```{r fig.width=8, fig.height=6, echo=FALSE}
protlabels <- limma_tab_wo6 %>%
      filter(logFC > 0 | logFC < 0,
             adj.P.Val < 0.05,
             Gene %in% c("ASAH1"))
                         #, "GPNMB", "SYNM", "CD14"))

size <- 0.8

@@

slim_data3 <- slim_data2 %>% 
  mutate(Abs_diff = rec-prim) %>%
  mutate(rank = row_number(Abs_diff)) %>% 
  dplyr::rename(Initial = prim, 
                Recurrent = rec) %>%
  ungroup() %>%
  filter(Protein %in% c("Q13510", 
                        "Q14956", 
                        "O15061",
                        "P05164")) %>%
  mutate(Gene = case_when(Protein == "Q13510" ~ "ASAH1",
                          Protein == "Q14956" ~ "GPNMB",
                          Protein == "O15061" ~ "SYNM",
                          Protein == "P05164" ~ "MPO"))

slim_data4 <- slim_data3 %>% 
  filter(Gene %in% c("ASAH1"))

### 2023-10-24 reports/html_gbm_recurrence_large_scale_proteolysis_manuscript_report.qmd
revision 2 commit

---
title: "GBM recurrence - Shot gun analysis of proteolytic products"
author: "Miguel Cosenza-Contreras"
format:
  gfm:
    toc: true 
    toc-depth: 4 # default is 3
    number-sections: true
    execute:
      output: true
always_allow_html: true
editor: source
---
```{r knirt_helper_funcs}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

@@

```{r}
#taking output from limma semi-specific
log2_semi_tryptic <- dplyr::select(compar_tab_peptwo6_annn2,
                                   logFC, 
                                   Protein, 
                                   peptide, 
                                   specificity, 
                                   is_terminal,
                                   Gene) %>%
  filter(specificity == "semi_specific",
         is_terminal == "not_terminal") %>%
  dplyr::rename(logFC_semi_peptide = logFC) %>%
  dplyr::select(-c(specificity, 
                   is_terminal))
```

@@

```{r}
log2semipept2_log2protein_spec <- left_join(log2_fully_tryptic, 
                                            log2_semi_tryptic) %>%
  na.omit() %>%
  mutate(DA_Protein = if_else(Protein %in% tab_sig_prots_tryptic$protein,
                              true = TRUE,
                              false = FALSE),
         DA_peptide = if_else(peptide %in% recwo6_pept22$peptide,
                              true = TRUE,
                              false = FALSE))
```

@@

```{r}
check_increased_pept <- cleavage_annoated_peptides %>%
  filter(Peptide %in% increased_recwo6_pept_annn)
```

@@

```{r}
cor.test(log2semipept2_log2protein_spec$logFC_fully_tryp_protein,
    log2semipept2_log2protein_spec$logFC_semi_peptide, 
    method = "pearson")
```
```{r}
ggsave(plot = scatter_proteolysis_plot, 
       filename = here::here("figures/scatter_proteolysis_plot_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 63)

@@

```{r}
increased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept_annn)

decreased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept_annn)

increased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept_annn) %>%
                    pull(cleave_area20)

decreased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept_annn) %>%
                    pull(cleave_area20)
```

```{r}
write_tsv(x = increased_Rec,
          file = here("results/semi_tryptic/proteolytic_products_increased_in_Recurrent.tsv"))

@@

@@

  write_rds(form_peptidesincreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_increasedREC.rds"))

@@

  write_rds(form_peptidesdecreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_decreasedREC.rds"))

@@

```{r}
write_tsv(x = diffabund_semis_fcs_annt,
          file = here("suppl_tables/Supplementary_Table_3_differential_abundant_semispepecific_peptides.tsv"))
```

@@

```{r}
write_tsv(x = gpnbm_supp_table,
          file = here("suppl_tables/Supplementary_Table_4_gpnbm_semispepecific_peptides.tsv"))
```

>>>>

---
title: "GBM recurrence - Shot gun analysis of proteolytic products"
author: "Miguel Cosenza-Contreras"
format:
  html: 
    toc-depth: 4 # default is 3
    toc-title: Contents
    toc-location: left # default is right, option is 'body'
    number-sections: true
    number-depth: 3
    embed-resources: true
always_allow_html: true
editor: source
---
```{r knirt_helper_funcs}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

@@

```{r}
#taking output from limma semi-specific
log2_semi_tryptic <- dplyr::select(compar_tab_peptwo6_annn,
                                   logFC, 
                                   protein, 
                                   peptide, 
                                   specificity, 
                                   is_terminal) %>%
  filter(specificity == "semi_specific",
         is_terminal == "not_terminal") %>%
  dplyr::rename(logFC_semi_peptide = logFC) %>%
  dplyr::select(-c(specificity, 
                   is_terminal))
```

@@

```{r}
log2semipept2_log2protein_spec <- left_join(log2_fully_tryptic, 
                                            log2_semi_tryptic)# %>%
  na.omit() %>%
  mutate(DA_Protein = if_else(Protein %in% tab_sig_prots_tryptic$protein,
                              true = TRUE,
                              false = FALSE),
         DA_peptide = if_else(peptide %in% recwo6_pept22$peptide,
                              true = TRUE,
                              false = FALSE))
```

@@

```{r}
check_increased_pept <- cleavage_annoated_peptides %>%
  filter(Peptide %in% increased_recwo6_pept)
```

@@

```{r}
cor.test(log2semipept2_log2protein_spec$logFC_fully_tryp_protein,
    log2semipept2_log2protein_spec$logFC_semi_peptide, 
    method = "pearson")
```

```{r}
ggsave(plot = scatter_proteolysis_plot, 
       filename = here::here("figures/scatter_proteolysis_plot_gbm.tiff"), 
       device = "tiff",
       units = "mm",
       width = 63,
       height = 63)

@@

```{r}
increased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept)

decreased_Rec <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept)

write_tsv(x = increased_Rec,
          file = here("results/semi_tryptic/proteolytic_products_increased_in_Recurrent.tsv"))

@@

increased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% increased_recwo6_pept) %>%
                    pull(cleave_area20)

decreased_Rec_4ice <- filter(cleave_areas$cleave_area20,
                            Peptide %in% decreased_recwo6_pept) %>%
                    pull(cleave_area20)
```

@@

  write_rds(form_peptidesincreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_increasedREC.rds"))
} else {

@@

  write_rds(form_peptidesdecreased_4ice, file = here("results/semi_tryptic/formated_pept_daglogo_decreasedREC.rds"))
} else {

@@

```{r}
write_tsv(x = diffabund_semis_fcs_annt,
          file = "suppl_tables/Supplementary_Table_3_differential_abundant_semispepecific_peptides.tsv")
```

@@

```{r}
write_tsv(x = gpnbm_supp_table,
          file = "suppl_tables/Supplementary_Table_4_gpnbm_semispepecific_peptides.tsv")
```
